Which ``Bad Smells'' Can Be Ignored?

### 2016-03-12 Manuscript.tex
final save for the fse submission

\definecolor{shadecolor}{gray}{0.9}
\conferenceinfo{ISCE}{'16 Austin, Texas}

\title{  Which ``Bad Smells'' can be Ignored?}
\numberofauthors{2} 
\author{  
\alignauthor
Rahul~Krishna, Tim Menzies  \\
       \affaddr{Computer Science, NC State, USA}\\
       {\{i.m.ralk,~tim.menzies\}@gmail.com}
\alignauthor
Lucas Layman \\
       \affaddr{Fraunhofer CESE, College Park, USA}\\ 
       {llayman@cese.fraunhofer.org}
\setlength{\columnsep}{7mm}}
\maketitle
\begin{abstract} 
The literature lists many   bad smells that may
be detrimental to a software system.
But which are relevant to the current project
and which can be ignored?
Developers, text books, and tools
disagree on which bad smells are important. 
Further, it is unclear at what threshold values smell-related metrics
should trigger code improvement or preventative maintenance. Most studies do not offer
{\em both} measurable metrics {\em and} a threshold for their proposed bad smells.

This paper proposes XTREE, a tool to prioritize  bad  smells.
Given an historical log of defects seen previously in the code, XTREE deprioritizes bad smells that have   no prior  association with defects.

@@

Based on these results, we endorse using bad smells to  
guiding refactoring. However,
not all bad smells
are relevant to the current project. It is hence important
to  prioritize the importance of specific bad smells in the context of each project.

@@

To address this problem, we propose the XTREE
tool to judge refactorings
proposed by developers.
  XTREE requires:
     \begin{itemize}
         \item An historical log of defects seen previously in the code;
         \item  Bad smells expressed in terms of the 
          metrics   in that log;
         \item Some proposed refactorings justified using those bad smells.
     \end{itemize}
Given that information,  XTREE can determine which
  proposed refactorings are be ignored
since they are not justifiable  
with respect to the history of the project.

@@

{\em PROBLEM \#3:}  
Shantnawi~\cite{Shatnawi10} warns most  studies do not offer
{\em both} a metric and a threshold for proposed bad smells. For example,  Darcy and Kemerer~\cite{darcy05}
discuss object-oriented (OO) design metrics that might defect bad smells (e.g. an excessive
depth of inheritance tree), but they do not define thresholds
for when the code starts to smell bad.  
Shatnami~\cite{Shatnawi10}
and  Alves et al.~\cite{Alves2010} are rare cases where both metrics and thresholds are defined (see \tion{methods}).
However, their approaches do not explore the interconnections between sets of code metrics. This is important if sets of
code metrics have to be changed at the same time.
For example, it is not possible to refactor a
long method it several smaller ones without increasing
the coupling metrics of a class. 
More generally, we say that bad smell thresholds should be
{\em conjunctions} of metric+threshold pairs that need to be changed {\em together}.
Shatnami~\cite{Shatnawi10}
and Alves et al. report {\em disjunctions} of multiple metrics+thresholds~\cite{Shatnawi10,Alves2010}
which could be separately and  inappropriately used
to guide refactoring.\footnote{From LL: I don't understand the conjunction/disjunction. Are you saying Metric1+Metric2 needs its own threshhold, or you need \{Metric1+Threshold1, Metric2+Threshold2\}. And what do Satnami and Alves do?}

@@

{\em PROBLEM \#4:} It is hard to judge the  effects of removing bad smells.
Refactored code cannot be assessed just by a rerun of the the test
suite since (refactorings are supposed to preserve behavior), and otherwise one must wait for new defect reports to evaluate the effect of refactoring changes on code.  
% waiting to observe 
% code refactorings is meant to
% preserve the behavior of the system (i.e. the same tests
% should pass before and after the refactoring). 
% Also, it many not be practical
%  to assessing refactorings by making the
% refactorings,  then seeing what happens to that code for the rest
% of the project. In a test-driven agile project, as
% SCRUM meetings change what stories are implemented next,
% the operational profile that lead to defects in the old
% code may not be repeated in future executions. 
To resolve this  problem, SE researchers such as 
Cheng et al.~\cite{Cheng10}, O'Keefe et al.~\cite{OKeeffe08,OKeeffe07},
Moghadam~\cite{Moghadam2011} and Mkaouer et al.~\cite{Mkaouer14}
use a pair of systems: 
\begin{enumerate}
    \item One system to propose changes;
\item A separate system to assess
how defective is the code before and after some
refactoring.
\end{enumerate}
For the second system,
 Cheng, O'Keefe, Moghadam and  Mkaouer et al. use the QMOOD hierarchical
quality model~\cite{Bansiya02}.
A shortcoming of QMOOD
is that quality models learned from other projects
may perform poorly when applied to new projects~\cite{me12d}.

@@

The rest of this paper is structured as follows. XXX.
discusses the cognitive
biases that make developers disagree on the relevance of those bad smells to
their particular projects.  
 That discussion is our motivation for developing the XTREE  tool
that   carefully assesses the merit of old bad smells to new projects.  XTREE is presented in Section 3 and evalauted in the rest of the paper.

XTREE   is focused on Fowler's bad smells
as contributors to software defects. Some researchers 
 suggest that discussing bad smells is useful for more just reducing defects. For example, Bosu and Carver~\cite{bosu13} report that
debates about code smells and refactorings can (a)~introduce new comers to the code base or (b)~share knowledge on a team
about interactions within the code. In that view, 
the concept of refactoring can be useful
even if it does not address issues of defects. 
That said, using
defect predictors,
it is possible to operationalize XTREE
while it might be harder to otherwise operationalize the assessment of bad smells.
Hence, 
this paper assumes that {\em better} bad smells and
bettering at finding code changes that  {\em reducing} the chances
that the new code will contain defects.

\subsection{Contributions of this Paper }
The XTREE learner has not been reported
previously.
 The experimental rig
 used to comparatively assess
    XTREE is another novel contribution. Using the design
    of that rig,
    other researchers could confirm/improve/refute
    the conclusions of this paper.
    Another contribution here is
    the comparative analysis of XTREE
    versus previously published approaches to learning
    bad smell detectors. that comparion shows
    that XTREE   rarely perfoms
      worse that the other approaches~\cite{Shatnawi10,Alves2010}
      and often is it   much better at proposing
      changes that reduce the expected
      value of the defects. Further, the number
    of code metrics mentioned by XTREE is far
    fewer than other approaches This means that XTREE is a better
    tool for ruling out   refactorings proposed by
    developers.

As to the relation to prior work, we
show here that XTREE is a  superior 
to the CD tool   developed previously with Borges~\cite{me12c}.

@@

Why build tools like XTREE to  critique proposed developer actions?
Are not those developers the best guide to what is important, or irrelevant
to their code?

It turns our that developer {\em cognitive biases} can mislead them into
asserting that some things are important and relevant when they are not.
This section discusses those biases-- firstly in the general sense
across all SE beliefs; then secondly in the   context of code smells.

A widespread malaise in software engineering is that
beliefs about software are rarely revised and hence may be
  inaccurate and 
misleading~\cite{passos11,jorgensen09,mei15,me16phase,prem16}. Software developers may hence have very
strong views on many issues, including bad smells, but those views may be wrong for the current
project.
It should be noted that  software engineering is not the only field with this problem.
  For example, the medical profession applies many practices based on studies that have been disproved (a recent article in the Mayo Clinic Proceedings~\cite{prasad13} found 146 medical practices based on studies in year i, but which were reversed by subsequent trials within years i + 10). Even when the evidence for or against a treatment or intervention is clear, medical providers and patients may not accept it~\cite{aschwanden10}. Aschwanden warns that {\em cognitive biases} such as confirmation bias (the tendency to look for evidence that supports what you already know and to ignore the rest) influence how we process information~\cite{aschwanden15}.

As in medicine, so too in software engineering.
According to Passos et al.~\cite{passos11},  developers often  assume that the lessons they learn from a few past projects are general to all their future projects. They comment ``past experiences were taken into account without much consideration for their context''~\cite{passos11}.  J{\o}rgensen \& Gruschke~\cite{jorgensen09} offer a similar warning. They report that the supposed software engineering ``gurus'' rarely use lessons from past projects to improve their future reasoning and that such poor past advice can be detrimental to new projects.~\cite{jorgensen09}.

To support the last paragraph, we offer several
specific studies
showing that   widely-held views in SE that are   now questionable, given recent results. 
The first few come from the general SE literature while the rest
relate specifically to bad smells:

\begin{enumerate} 
\item
Nagappan et al. analyzed~\cite{mei15} 
how   {\tt goto}  was used  in 11,000 Github repositories. 
They found that developers
rarely used {\tt goto} in the unrestricted manner that alarmed Dijkstra in his  1968
memo ``Go to statement considered harmful''~\cite{Dijkstra68}.
\item
In other work~\cite{me16phase} with the Software Engineering Institute (SEI), we have revisited
the ``phase delay'' truism that {\em fixing a requirements error  during
coding can be thousands of times more expensive than recognizing and fixing that error
as soon as it arises}. 
We found no evidence for such large   phase delay effect in modern
software (in 171 software projects shepherded
by the SEI, 2006-2014) possibly since it has been  heavily mitigated by $21^{st}$ software development languages, tools and development practices. 
Whatever the reason, the main point  is that phase delay is a widely
held belief, despite little recent evidence to support it.
\item
Devanbu et al.  examined responses from 564 Microsoft software developers from around
the world, They found that  ``(a)~programmers do indeed have very
strong beliefs on certain topics; (b)~their beliefs are primarily formed
based on personal experience, rather than on findings in empirical
research; (c)~beliefs can vary with each project, but do not necessarily
correspond with actual evidence in that project''~\cite{prem16}.
\end{enumerate}
Devanbu et al. further  comment that ``programmers give personal experience
as the strongest influence in forming their opinions''. This is a troubling
result, especially given the above comments from Passos,  J{\o}rgensen et al.~\cite{passos11,jorgensen09} about how quickly practitioners form, then freeze, those opinions for the rest of their
careers.

@@

If the above remarks hold true for bad smells, then we would expect
to see much disagreement on what bad smells are important and relevant
to  a particular project. This is indeed the case.
The first column of \fig{smells} 
lists  commonly mentioned bad smells and comes from Fowler's 1999 text~\cite{Fowler99} and a subsequent 2005 text by Kerievsky that is widely cited~\cite{Kerievsky2005}.
The other
columns show other work that comments on which bad smells matter most.
The columns marked as Lanza'06, Yamashita'13, and Atwood'06 come from
peer reviewed literature. The column maked SonarQube is an open source
code assessment tool that includes dectectors for six of the bad smells
in column one. 
The {\em developer survey}   (in the right-hand-side column),
   this shows the results of a hour-long whiteboard sessions with a group of a
 dozen developers from one Washington
    D.C. web-tools development organization. In that study, participants
    worked in a round robin manner to rank the bad smells they thought were
    important (and any disagreements were discussed with the whole group).
     Amongst those group, there was there was  some
    consensus on  the priority of fixing which  bad smells   
    (see the annotations VH=very high,
    H=high, L=low, VL=very low, and ``?''= no consensus).

@@

A  blank cell in \fig{smells}
indicates where   other work has chosen to ignore
one the   bad smells in column one. 
Note that most of the cells are blank.
While the Atwood'06~\cite{Atwood06} study  comments on   
all the   bad smells,
all the  other  studies ignore the majority of the Fowler bad smells.
For exampel SonarQube has no detectors for many of the column one bad smells.
Also, nearly half the Yamashita list of bad smells
    does not appear in the Fowler and Kerievsky list
    (the eight numbers
    in the  Yamashita'13 column show the rankings for the bad smells 
    that overlap with Fowler and Kerievsky).

Two of the studies in \fig{smells} offers some comments on the relative importance
of the different bad smells the   Yamashita'13 study and the developer study.Three of the bad smells listed in the top half of the Yamashita'13 rankings also score very high in the developer survey. Those three were {\em duplicated code, large class}, 
        and {\em long method}. 
 Note that this agreement also means that the
  Yamashita'13 study and the {\em developer survey}   
  believe
        that very few  code smells are   high priority issues
        requiring refactoring.

In summary, just because a developer strongly believes something
about bad smells, does
not necessary mean that that belief is useful for the current project.
Human developers can be clever, but their thinking can also be distorted
by cognitive biases.
Hence, as shown in \fig{smells}, developers and text books and tools 
can disagree on which bad smells are important.
Special tools are needed to assess their beliefs in, for example,
bad smells.

\section{Learning Bad Smell Thresholds}
The SE literature offers  two ways to 
learning bad smell thresholds.
One way relies on 
{\em outlier statistics}~\cite{erni96,bender99}. This way
has been used   by Shatnawi~\cite{Shatnawi10}, Alvenes et al.~\cite{Alves2010}
    and Hermans et al.~\cite{hermans15}.
    Another approach is 
    based on {\em cluster deltas} that we developed
    for   Centroid Deltas~\cite{me12c} and 
    use here for XTREE.

These two approaches are discussed below.
\subsection{  Outliers}\label{sect:p}

The outlier approach assume that unusually large measurements are risk-prone.
Hence, they generate one bad smell threshold for any metric
with such an ``unusually large'' measurements. 
The literature lists several ways to define ``unusually large''.

\subsubsection{Enri \& Lewerentz}
One of the first methods was proposed Erni and Lewerentz~\cite{erni96}. 
Given classes described with the  code metrics of \fig{ck},
Enri and Lewerentz found the   mean $\mu$ and the standard deviation $\sigma$
of each
code metrics. Their definition of problematic outlier was any code
metric with a measurement greater than $\mu+\sigma$.

Shatnami and Alves et al.~\cite{Shatnawi10,Alves2010} depreciate
using $\mu+\sigma$. since it does not   consider the fault-proneness of classes when the thresholds are computed. Also, the method lacks  empirical validation.

\subsubsection{ Shatnami}
Shatnami~\cite{Shatnawi10}'s preferred alternative to $\mu+\sigma$
is to use the VARL method (Value of Acceptable Risk Level) proposed by Bender~\cite{bender99} in his epidemiology studies.  This approach uses two
thresholds which, following Shatnami's guidance, we set to
$p_0=p_1=0.05$.

VARL encodes the defect count
for each class to 0 (for no defects known) to 1 (for defects known).
Univariate binary logistic regression is applied to learn three coefficients:  
 $\alpha$ is  constat intercept;
    $\beta$ is the coefficient for maximizing log-likelihood;
  and $p_0$  
    measures   how well this   model predicts for   defects.
Any code metric with $p_0>0.05$ is  rejected as being a poor defect predictor. Thresholds are then learned from the surviving metrics $M_c$ using
an equation proposed by Bender:
\begin{equation}
 \mathit{bad\; smell\; if\;} M_c > \frac{1}{\beta }\left( {\log \left( {\frac{{{p_1}}}{{1 - {p_1}}}} \right) - \alpha } \right) 
\end{equation}
\subsubsection{ Alves et al.}
Alves et al.~\cite{Alves2010} propose another approach
to finding thresholds that  utilizes the underlying statistical distribution and scale of the metrics. 
Evey metric value is weighted according to the source lines of code (LOC) of the class. All the weighted metrics are then normalized; i.e. they are divided by the sum of all weights of the same system. For consistency
with Shatnawi, we explore the correlation between
the code metrics and the defect counts and   reject code metrics that are poor predictors of defects (i.e.   those  with $p > 0.05$).
Following this, the normalized metric values are ordered in an ascending fashion (this is equivalent to computing a density function, in which the x-axis represents the weight ratio (0-100\%), and the y-axis the metric scale).
Alves et al. then compute their 
thresholds   by choosing the percentage of the overall code that needs to be represented. For instance, Alves et al. suggest the use 90\% quantile of the overall code to derive the threshold for a specific metric. The intuition here is that the worst
code have outliers beyond 90\% of the normal code measurements.

Our reading of the literature is that the Alves et al.  
approach is preferred to that of Shatnami. For example, Herman et al.~\cite{hermans15} used the
Alves et al. method in their  2015 paper on
exploring bad smells.

@@

\subsubsection{Discussion }\label{sect:disc}
The advantage of the outlier
approach is that it is simple to implement. 
For example, in less than a day we could
code up and apply  the Shatnami and 
Alves et al. methods.

The outlier approaches has   two  major disadvantages. 
Firstly, they are {\em verbose}; i.e.  they can
propose thresholds on most of the code metrics.
Such verbosity makes it hard   to critique and prune bad smells
since any bad smell mentioned by a developer has some support
somewhere amongst all those proposed thresholds.

Secondly, the outlier approaches suffers from
the {\em disjunction} problem discussed in the introduction.
That is, while
 they propose thresholds for many code metrics
individually, they make no comment on what
metrics need to be changed at the same time.

@@

Cluster deltas are a general method
for learning {\em conjunctions} of changes
that need to be applied at the same time. 
This approach works as follows:
\begin{itemize}
    \item Cluster the data. 
    \item Find
neighboring clusters $C_+,C_-$ where $C_+$ has more examples of defective
modules than $C_-$;
\item Compute the  delta   in code metrics between the clusters using \mbox{$\Delta = C_+ - C_-$};
\item $\Delta$ are changes needed in defective modules of $C_+$ to
      make them more like the less-defective modules of $C_-$
\end{itemize}
Note that $\Delta$ is a conjunction of  recommendations.
Since it is computed
from neighboring clusters, it   reasons about
examples with similar distributions.  Hence,   $\Delta$  respects the naturally occurring constraints in the data. For example,
repeating the above example about large methods,   $\Delta$   will not  suggest lowering lines of code
without also increasing a coupling measure

Cluster deltas are used in CD~\cite{me12c} and XTREE.

@@

As shown below, one drawback with with CD is that it is very {\em verbose}
since
CD   recommended changes to all code
metrics with different values in those two centroids. 
This makes it hard to use CD to   critique and prune away bad smells.

Another drawback with CD is revealed by the comparison experiments
in this paper: CD never worked as well as XTREE in proposing
changes to reduce defects.

@@

  XTREE  is a cluster delta algorithm
  that avoids the problem of verbose $\Delta$s.
  Instead of reasoning over cluster centroids,
  XTREE utilizes a decision tree learning approach
  to find the fewest differences groupings of examples.

  Decision trees are built via
  {\em iterative dichotomization} that  finds  a  split  in  the  values  of  independent  features  that  most  reduces the variability of the dependent feature. The data is then divided on that split
  and iterative dichotomization is called
  recursively on the data in each split. In this way,
  each split is the root of a sub-tree.

>>>>

\definecolor{shadecolor}{gray}{0.9}
\conferenceinfo{FSE}{'16 Seattle, Washington}

\title{  Which ``Bad Smells'' Can Be Ignored?}
\numberofauthors{2} 
\author{  
\alignauthor
Rahul~Krishna, Tim Menzies  \\
       \affaddr{Computer Science, NC State, USA}\\
       {\{i.m.ralk,~tim.menzies\}@gmail.com}
\alignauthor
Lucas Layman \\
       \affaddr{Fraunhofer CESE, College Park, USA}\\ 
       {llayman@cese.fraunhofer.org}
\setlength{\columnsep}{7mm}}
\maketitle
\begin{abstract} 
The literature lists many   bad smells that may
be detrimental to a software system.
But which are relevant to the current project
and which can be ignored?
Developers, text books, and tools
disagree on which bad smells are important. 
Further, it is unclear at what threshold values smell-related metrics
should trigger code improvement or preventative maintenance. Most studies do not offer
{\em both} measurable metrics {\em and} a threshold for their proposed bad smells.

This paper proposes XTREE, a tool to prioritize  bad  smells.
Given a historical log of defects seen previously in the code, XTREE deprioritizes bad smells that have   no prior  association with defects.

@@

It is hence important to  prioritize the importance of specific bad smells in the context of each project since not all bad smells are relevant to the current project. Based on these results, we endorse using bad smells to guiding refactoring.

@@

To address this problem, we propose the XTREE
tool to judge refactorings
proposed by developers.
  XTREE requires:
     \begin{itemize}
         \item A historical log of defects seen previously in the code;
         \item  Bad smells expressed in terms of the 
          metrics   in that log;
         \item Some proposed refactorings justified using those bad smells.
     \end{itemize}
Given that information,  XTREE can determine which
  proposed refactorings are to be ignored
since they are not justifiable  
with respect to the history of the project.

@@

{\em PROBLEM \#3:}  
Shantnawi~\cite{Shatnawi10} warns most  studies do not offer
{\em both} a metric and a threshold for proposed bad smells. For example,  Darcy and Kemerer~\cite{darcy05}
discuss object-oriented (OO) design metrics that might defect bad smells (e.g. an excessive
depth of inheritance tree), but they do not define thresholds
for when the code starts to smell bad.  
Shatnawi~\cite{Shatnawi10}
and  Alves et al.~\cite{Alves2010} are rare cases where both metrics and thresholds are defined (see \tion{bst}).
However, their approaches do not explore the interconnections between sets of code metrics. This is important if sets of
code metrics have to be changed at the same time.
For example, it is not possible to refactor a
long method if several smaller ones without increasing
the coupling metrics of a class. 
More generally, we say that bad smell thresholds should be
{\em conjunctions} of metric+threshold pairs that need to be changed {\em together}.
Shatnawi~\cite{Shatnawi10}
and Alves et al. report {\em disjunctions} of multiple metrics+thresholds~\cite{Shatnawi10,Alves2010} without indicating how changes to these metrics are related.

@@

{\em PROBLEM \#4:} It is hard to judge the  effects of removing bad smells.
Refactored code cannot be assessed just by a rerun of the the test
suite since (refactorings are supposed to preserve behavior), and otherwise one must wait for new defect reports to evaluate the effect of refactoring changes on code.  
% waiting to observe 
% code refactorings is meant to
% preserve the behavior of the system (i.e. the same tests
% should pass before and after the refactoring). 
% Also, it many not be practical
%  to assessing refactorings by making the
% refactorings,  then seeing what happens to that code for the rest
% of the project. In a test-driven agile project, as
% SCRUM meetings change what stories are implemented next,
% the operational profile that lead to defects in the old
% code may not be repeated in future executions. 
To resolve this  problem, SE researchers such as 
Cheng et al.~\cite{Cheng10}, O'Keefe et al.~\cite{OKeeffe08,OKeeffe07},
Moghadam~\cite{Moghadam2011} and Mkaouer et al.~\cite{Mkaouer14}
use a pair of systems: 
\begin{enumerate}
    \item One system to propose changes;
\item A separate system to assess
how defective is the code before and after some
refactoring.
\end{enumerate}
For the second system,
 Cheng, O'Keefe, Moghadam and  Mkaouer et al. use the QMOOD hierarchical
quality model~\cite{Bansiya02}.
A shortcoming of QMOOD
is that quality models learned from other projects
may perform poorly when applied to new projects~\cite{localvsglobal}.

@@

The rest of this paper is structured as follows. 
The next section
discusses the cognitive
biases that make developers disagree on the relevance of those bad smells to
their particular projects
(that discussion is our motivation for developing 
tools like XTREE). This is followed by a discussion
on various methods to learn thresholds for bad smells.
After evaluating those methods, we discuss various threats to
validity that threaten our conclusions.

Before  that, we offer one following note on a
core assumption of XTREE.
XTREE   is focused on Fowler's bad smells
as contributors to software defects. Some researchers 
 suggest that discussing bad smells is useful for more just reducing defects. For example, Bosu and Carver~\cite{bosu13} report that
debates about code smells and refactorings can (a)~introduce new comers to the code base or (b)~share knowledge on a team
about interactions within the code. In that view, 
the concept of refactoring can be useful
even if it does not address issues of defects. 
% That said, using
% defect predictors,
% it is possible to operationalize XTREE
% while it might be harder to otherwise operationalize the assessment of bad smells.
% Hence, 
% this paper assumes that {\em better} bad smells and
% bettering at finding code changes that  {\em reducing} the chances
% that the new code will contain defects.

\subsection{Contributions of this Paper }
The XTREE learner has not been reported
previously. 
The comparative analysis of XTREE
    versus previously published approaches to learning
    bad smell detectors shows
    that XTREE   rarely perfoms
      worse that the other approaches~\cite{Shatnawi10,Alves2010}
      and often is it   much better at proposing
      changes that reduce the expected
      number of of the defects. Further, the number
    of code metrics mentioned by XTREE is far
    fewer than other approaches, making it easier to validate refactorings proposed by
    developers. 
We
show here that XTREE is less verbose and offers better recommendations than the CD tool   developed previously with Borges~\cite{me12c}.
 The experimental rig
 used to comparatively assess
    XTREE is another novel contribution\footnote{tiny.cc/XTREE}. Using the design
    of that rig,
    other researchers could confirm/improve/refute
    the conclusions of this paper.

@@

Why build tools like XTREE to  critique proposed developer actions?
Are not those developers the best guide to what is important or irrelevant
to their code? Developer {\em cognitive biases} can mislead them to
assert that some things are important and relevant when they are not.
This section discusses those biases-- firstly in the general sense
across all SE beliefs; then secondly in the   context of code smells.

A widespread malaise in software engineering is that
beliefs about software are rarely revised and hence may be
  inaccurate and 
misleading~\cite{passos11,jorgensen09,mei15,me16phase,prem16}. Software developers may have strong views on many issues, including bad smells, but those views may be wrong for the current
project.
Software engineering is not the only field with this problem.
  For example, the medical profession applies many practices based on studies that have been disproved (a recent article in the Mayo Clinic Proceedings~\cite{prasad13} found 146 medical practices based on studies in year i, but which were reversed by subsequent trials within years i + 10). Even when the evidence for or against a treatment or intervention is clear, medical providers and patients may not accept it~\cite{aschwanden10}. Aschwanden warns that {\em cognitive biases} such as confirmation bias (the tendency to look for evidence that supports what you already know and to ignore the rest) influence how we process information~\cite{aschwanden15}.

As in medicine, so too in software engineering.
According to Passos et al.~\cite{passos11},  developers often  assume that the lessons they learn from a few past projects are general to all their future projects. They comment ``past experiences were taken into account without much consideration for their context''~\cite{passos11}.  J{\o}rgensen \& Gruschke~\cite{jorgensen09} offer a similar warning. They report that the supposed software engineering ``gurus'' rarely use lessons from past projects to improve their future reasoning and that such poor past advice can be detrimental to new projects.~\cite{jorgensen09}. Other studies have shown some widely-held views are   now questionable given new evidence:

\begin{enumerate} 
% COMMENT FROM LL: This bullet is weak. Maybe goto isn't used in an unrestricted manner because of Djikstra's warning.
% \item
% Nagappan et al. analyzed~\cite{mei15} 
% how   {\tt goto}  was used  in 11,000 Github repositories. 
% They found that developers
% rarely used {\tt goto} in the unrestricted manner that alarmed Dijkstra in his  1968
% memo ``Go to statement considered harmful''~\cite{Dijkstra68}.
\item
In other work~\cite{me16phase} with the Software Engineering Institute (SEI), we have revisited
the ``phase delay'' truism that {\em the cost of fixing an error dramatically increases the longer it is in the system}. 
We found no evidence for such large   phase delay effect in modern
software (in 171 software projects shepherded
by the SEI, 2006-2014) possibly since it has been  heavily mitigated by $21^{st}$ century software development languages, tools and development practices. 
Whatever the reason, the main point  is that phase delay is a widely
held belief, despite little recent evidence to support it.
\item
Devanbu et al.  examined responses from 564 Microsoft software developers from around
the world, they found that  ``(a)~programmers do indeed have very
strong beliefs on certain topics; (b)~their beliefs are primarily formed
based on personal experience, rather than on findings in empirical
research; (c)~beliefs can vary with each project, but do not necessarily
correspond with actual evidence in that project''~\cite{prem16}.
\end{enumerate}
Devanbu et al. further  comment that ``programmers give personal experience
as the strongest influence in forming their opinions''. This is a troubling
result, especially given the above comments from Passos and  J{\o}rgensen et al.~\cite{passos11,jorgensen09} about how quickly practitioners form, freeze, and rarely revist those opinions.

@@

If the above remarks hold true for bad smells, then we would expect
to see much disagreement on which bad smells are important and relevant
to  a particular project. This is indeed the case.
The first column of \fig{smells} 
lists  commonly mentioned bad smells and comes from Fowler's 1999 text~\cite{fowler99} and a subsequent 2005 text by Kerievsky that is widely cited~\cite{Kerievsky2005}.
The other
columns show data from other studies on which bad smells matter most.
The columns marked as Lanza'06 and Yamashita'13 are from peer reviewed literature. The column maked SonarQube is a popular open source
code assessment tool that includes detectors for six of the bad smells
in column one. 
The {\em developer survey}   (in the right-hand-side column),
   this shows the results of an hour-long whiteboard sessions with a group of 12 developers from a Washington
    D.C. web tools development company. Participants
    worked in a round robin manner to rank the bad smells they thought were
    important (and any disagreements were discussed with the whole group).
     Amongst the group, there was  some
    consensus on  the priority of which bad smells to fix
    (see the annotations VH=very high,
    H=high, L=low, VL=very low, and ``?''= no consensus).

@@

A  blank cell in \fig{smells}
indicates where   other work has chosen to ignore
one the   bad smells in column one. 
Note that most of the cells are blank, and that the studies ignore the majority of the Fowler bad smells.
SonarQube has no detectors for many of the column one bad smells.
Also, nearly half the Yamashita list of bad smells
    does not appear in the Fowler and Kerievsky list. The eight numbers
    in the  Yamashita'13 column show the rankings for the bad smells 
    that overlap with Fowler and Kerievsky; Yamashita also discussed other smells not covered in Fowler'99.

Two of the studies in \fig{smells} offers some comments on the relative importance
of the different bad smells in the Yamashita'13 study and the developer study. Three of the bad smells listed in the top half of the Yamashita'13 rankings also score very high in the developer survey. Those three were {\em duplicated code, large class}, 
        and {\em long method}. 
 Note that this agreement also means that the
  Yamashita'13 study and the developer survey   
  believe
        that very few  code smells are   high priority issues
        requiring refactoring.

In summary, just because one developer strongly believes in the importance of a bad smells does not mean that belief transfers to other developers or projects.
Developers can be clever, but their thinking can also be distorted
by cognitive biases.
Hence, as shown in \fig{smells}, developers, text books, and tools 
can disagree on which bad smells are important.
Special tools are needed to assess their beliefs, for example, their beliefs in
bad smells.

\section{Learning Bad Smell Thresholds}\label{sect:bst}
Research studies and tools (e.g., SonarQube~\cite{sq15}) leverage OO metrics as representatives for code smells, as we do in this study. In this section, we focus on determining thresholds for interpreting these metrics as ``bad smells'' or not.

The SE literature offers  two ways to 
learning bad smell thresholds.
One approach relies on 
{\em outlier statistics}~\cite{erni96,bender99}. This approach
has been used   by Shatnawi~\cite{Shatnawi10}, Alvenes et al.~\cite{Alves2010}
    and Hermans et al.~\cite{hermans15}.
    Another approach is 
    based on {\em cluster deltas} that we developed
    for   Centroid Deltas~\cite{me12c} and 
    use here for XTREE. 
    These two approaches are discussed below.
\subsection{  Outlier Statistics}\label{sect:p}

The outlier approach assume that unusually large measurements indicate risk-prone code.
Hence, they generate one bad smell threshold for any metric
with such an ``unusually large'' measurements. 
The literature lists several ways to define ``unusually large''.

\subsubsection{Enri \& Lewerentz}
Given classes described with the  code metrics of \fig{ck},
Enri and Lewerentzz~\cite{erni96} found the   mean $\mu$ and the standard deviation $\sigma$
of each
code metrics. Their definition of problematic outlier was any code
metric with a measurement greater than $\mu+\sigma$.
Shatnawi and Alves et al.~\cite{Shatnawi10,Alves2010} depreciate
using $\mu+\sigma$ since it does not consider the fault-proneness of classes when the thresholds are computed. Also, the method lacks  empirical validation.

\subsubsection{ Shatnawi}
Shatnawi~\cite{Shatnawi10}'s preferred alternative to $\mu+\sigma$
is to use the VARL method (Value of Acceptable Risk Level) proposed by Bender~\cite{bender99} in his epidemiology studies.  This approach uses two
thresholds which, following Shatnawi's guidance, we set to
$p_0=p_1=0.05$.

VARL encodes the defect count
for each class as 0 (no defects known in class) or 1 (defects known in class).
Univariate binary logistic regression is applied to learn three coefficients:  
 $\alpha$ is the intercept constant;
    $\beta$ is the coefficient for maximizing log-likelihood;
  and $p_0$  
    measures   how well this   model predicts for   defects.
Any code metric with $p_0>0.05$ is  rejected as being a poor defect predictor. Thresholds are then learned from the surviving metrics $M_c$ using
an equation proposed by Bender:
\begin{equation}
 \mathit{bad\; smell\; if\;} M_c > \frac{1}{\beta }\left( {\log \left( {\frac{{{p_1}}}{{1 - {p_1}}}} \right) - \alpha } \right) 
\end{equation}
For consistency
with Shatnawi, we explore the correlation between
the code metrics and the defect counts and   reject code metrics that are poor predictors of defects (i.e.   those  with $p > 0.05$).
\subsubsection{ Alves et al.}
Alves et al.~\cite{Alves2010} propose another approach
to finding thresholds that  utilizes the underlying statistical distribution and scale of the metrics. 
Metric values for each class are weighted according to the source lines of code (SLOC) of the class. All the weighted metrics are then normalized by the sum of all weights of for the system. 
The normalized metric values are ordered in an ascending fashion (this is equivalent to computing a density function, in which the x-axis represents the weight ratio (0-100\%), and the y-axis the metric scale).
Alves et al. then select a percentage value (they suggest 90\%) which represents the ``normal'' values for metrics. The metric threshold, then, is the metric value for which 90\% of the classes fall below. The intuition  is that the worst code has outliers beyond 90\% of the normal code measurements. Hermans et al.~\cite{hermans15} used the
Alves et al. method in their  2015 paper on
exploring bad smells.

@@

\subsubsection{Discussion }\label{sect:disc}
The advantage of the outlier-based
approaches is that they are simple to implement, but the approaches have   two  major disadvantages. 
First, they are {\em verbose}. A threshold can be calculated for every metric -- so, which one should the developers focus on changing? Without a means for prioritizing the  thresholds and metrics against one another, developers may have numerous or conflicting recommendations on what to improve. Second, the outlier approaches suffers from
the {\em disjunction} problem.
That is, while
 they propose thresholds for many code metrics
individually, they make no comment on what
metrics need to be changed at the same time.

@@

Cluster deltas are a general method
for learning {\em conjunctions} of changes
that need to be applied at the same time. 
This approach works as follows:
\begin{itemize}
    \item Cluster the data. 
    \item Find
neighboring clusters $C_+,C_-$ where $C_+$ has more examples of defective
modules than $C_-$;
\item Compute the  delta   in code metrics between the clusters using \mbox{$\Delta = C_+ - C_-$}, i.e.
{\em towards} the cluster with lower defects;
\item $\Delta$ are changes needed in defective modules of $C_+$ to
      make them more like the less-defective modules of $C_-$
\end{itemize}
Note that $\Delta$ is a conjunction of  recommendations.
Since it is computed
from neighboring clusters, the examples contain similar distributions and $\Delta$ respects the naturally occurring constraints in the data. For example,
given a bad smell pertaining to large methods,   $\Delta$   will not  suggest lowering lines of code
without also increasing a coupling measure. 
Cluster deltas are used in CD~\cite{me12c} and XTREE.

@@

As shown below, one drawback with with CD is that it is {\em verbose}
since
CD   recommended changes to all code
metrics with different values in those two centroids. 
This makes it hard to use CD to   critique and prune away bad smells. Further, CD will be shown to be
not as effective
in proposing changes to reduce defects as XTREE.

@@

  XTREE  is a cluster delta algorithm
  that avoids the problem of verbose $\Delta$s.
  Instead of reasoning over cluster centroids,
  XTREE utilizes a decision tree learning approach
  to find the fewest differences between clusters of examples.

 XTREE uses an iterative dichotomizer that
first converts the values for each code
metrics into a small number of ranges.
Consider a code metric is split into $r$ ranges and each range is of
  size
    $n_r$ is
  associated with a set of defect counts $x$ with standard deviation
  $\sigma_r$.
  The best split for that range is
  the one that minimizes the expected value of the
  defect variance, after the split; i.e.
  $\sum_r\frac{n_r}{n}\sigma_x$ (where $n=\sum_r n_r$).
  This discretizer then recurses on each part of the split
  to find other splits.
